---
title: "t vs Z Distribution"
author: "Rick"
date: "February 2, 2017"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## t vs Z Distribution
The t distribution, invented by William Gosset in 1908, has thicker tails than the normal. Also, instead of having two parameters, mean and variance, as the normal does, the t distribution has only one - the number of degrees of freedom (df). 

As df increases, the t distribution gets more like a standard normal, so it's centered around 0. Also, the t assumes that the underlying data are iid Gaussian so the statistic (X' - mu)/(s/sqrt(n)) has n-1 degrees of freedom. 

## R code for investigating t and Z densities
```{r echo = TRUE, eval = FALSE}
library(ggplot2)
k <- 1000
xvals <- seq(-5, 5, length = k)
myplot <- function(df){
  d <- data.frame(y = c(dnorm(xvals), dt(xvals, df)),
                  x = xvals,
                  dist = factor(rep(c("Normal", "T"), c(k,k))))
  g <- ggplot(d, aes(x = x, y = y)) 
  g <- g + geom_line(size = 2, aes(colour = dist))
  g <- g + ggtitle(paste("df=", df, ", dnorm(0) - dt(0, df)="
                         , round(dnorm(0) - dt(0, df), 3)))
  print(g)
}
myplot(3)
myplot(20)
```
https://github.com/swirldev/swirl_courses/blob/master/Statistical_Inference/T_Confidence_Intervals/tPlot.R


## Plots for investigating t and Z densities
```{r fig.height=3}
library(ggplot2)
k <- 1000
xvals <- seq(-5, 5, length = k)
myplot <- function(df){
  d <- data.frame(y = c(dnorm(xvals), dt(xvals, df)),
                  x = xvals,
                  dist = factor(rep(c("Normal", "T"), c(k,k))))
  g <- ggplot(d, aes(x = x, y = y)) 
  g <- g + geom_line(size = 2, aes(colour = dist))
  g <- g + ggtitle(paste("df=", df, ", dnorm(0) - dt(0, df)="
                         , round(dnorm(0) - dt(0, df), 3)))
  print(g)
}
myplot(3)
myplot(20)
```


## R code for investigating the upper quantiles of the t and Z densities
```{r echo = TRUE, eval = FALSE}
library(ggplot2)
pvals <- seq(.5, .99, by = .01)
myplot2 <- function(df){
  d <- data.frame(n= qnorm(pvals),t=qt(pvals, df),
                  p = pvals)
  g <- ggplot(d, aes(x= n, y = t))
  g <- g + geom_abline(size = 2, col = "lightblue")
  g <- g + geom_line(size = 2, col = "black")
  g <- g + geom_vline(xintercept = qnorm(0.975))
  g <- g + geom_hline(yintercept = qt(0.975, df))
  g <- g + ggtitle(paste("df=", df, ", qt(0.975, df) - qnorm(0.975)="
                         , round(qt(0.975, df) - qnorm(0.975), 3)))
  print(g)
}
myplot2(3)
myplot2(20)
```

https://github.com/swirldev/swirl_courses/blob/master/Statistical_Inference/T_Confidence_Intervals/tQuant.R

## Plots for investigating the upper quantiles of the t and Z densities
```{r fig.height=3}
library(ggplot2)
pvals <- seq(.5, .99, by = .01)
myplot2 <- function(df){
  d <- data.frame(n= qnorm(pvals),t=qt(pvals, df),
                  p = pvals)
  g <- ggplot(d, aes(x= n, y = t))
  g <- g + geom_abline(size = 2, col = "lightblue")
  g <- g + geom_line(size = 2, col = "black")
  g <- g + geom_vline(xintercept = qnorm(0.975))
  g <- g + geom_hline(yintercept = qt(0.975, df))
  g <- g + ggtitle(paste("df=", df, ", qt(0.975, df) - qnorm(0.975)="
                         , round(qt(0.975, df) - qnorm(0.975), 3)))
  print(g)
}
myplot2(3)
myplot2(20)
```

